---
title: "Resampling"
author: "Michael Raminski"
date: "May 9, 2017"
output: html_document
---

**KNIT YOUR DOCUMENT AS *HTML* AND SUBMIT IT AND THE `Rmd` file**

## Resampling

`x` is a random variable. We want to not only know what the `mean(x)` is but want to calculate the uncertainty of `mean(x)`.  Measuring the uncertainty requires repeated measurements of `mean(x)`.

- Calculate the mean of `x`.
- Calculte the `sd( mean(x) )` using the **using 10-fold resampling**.  Create your own folds, show your work.
- Calculte the `sd( mean(x) )` using the **using 10 bootstrap sample**.  Create your own folds, show your work. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(dplyr)
library(tibble)
library(tibble)
library(Hmisc)
library(pastecs)
library(reshape2)
library(ggplot2)
library(ggm)
library(corrgram)
library(psych)
``` 

###Generate 20 random numbers
```{r, warning=FALSE, message=FALSE, error=FALSE, eval=TRUE}
set.seed(1) 
x <- runif(20,1,20)
```

###Mean of 20 random numbers
```{r, warning=FALSE, message=FALSE, error=FALSE, eval=TRUE}
mean(x)
```

###Standard Deviation of mean(x) using 10-fold resampling 
```{r, warning=FALSE, message=FALSE, error=FALSE, eval=TRUE}
df_cv <- data.frame(mean(x[-1:-2]),mean(x[-3:-4]),mean(x[-5:-6]),mean(x[-7:-8]),mean(x[-9:-10]),mean(x[-11:-12]),mean(x[-13:-14]),mean(x[-15:-16]),mean(x[-17:-18]),mean(x[-19:-20]))

sd_cv <- sd(df_cv)
```

###Standard Deviation of mean(x) using 10 bootstrap sample
```{r, warning=FALSE, message=FALSE, error=FALSE, eval=TRUE}
df_boot <- data.frame(mean(sample(x, 20, replace = TRUE)),mean(sample(x, 20, replace = TRUE)),mean(sample(x, 20, replace = TRUE)),mean(sample(x, 20, replace = TRUE)),mean(sample(x, 20, replace = TRUE)),mean(sample(x, 20, replace = TRUE)),mean(sample(x, 20, replace = TRUE)),mean(sample(x, 20, replace = TRUE)),mean(sample(x, 20, replace = TRUE)),mean(sample(x, 20, replace = TRUE)))

sd_boot <- sd(df_boot)
```


### Your answers here: 

- sd_cv   is: `r sd_cv`
- sd_boot is: `r sd_boot`
