---
title: "Classification Mettrics"
author: "Michael Raminski"
date: "May 7, 2017"
output: html_document
---


**KNIT YOUR DOCUMENT AS *HTML* AND SUBMIT IT AND THE `Rmd` file.** 
 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(dplyr)
library(tibble)
library(tibble)
library(Hmisc)
library(pastecs)
library(reshape2)
library(ggplot2)
library(ggm)
library(corrgram)
library(psych)
``` 

## Classification Metrics Functions 
Write functions of `y` and `y_hat` to calculate the following classification metrics


####True Positive Rate
```{r, warning=FALSE, message=FALSE, error=FALSE, eval=TRUE}
True_Positive_Rate <- function(y,y_hat){
  return(sum(y == 1 & y_hat == 1, na.rm=TRUE) / sum(y == 1 & y_hat != "NA", na.rm=TRUE))}
```

####False Positive Rate
```{r, warning=FALSE, message=FALSE, error=FALSE, eval=TRUE}
False_Positive_Rate <- function(y,y_hat){
  return(sum(y == 0 & y_hat == 1, na.rm=TRUE) / sum(y == 0 & y_hat != "NA", na.rm=TRUE))}
```

####True Negative Rate
```{r, warning=FALSE, message=FALSE, error=FALSE, eval=TRUE}
True_Negative_Rate <- function(y,y_hat){
  return(sum(y == 0 & y_hat == 0, na.rm=TRUE) / sum(y == 0 & y_hat != "NA", na.rm=TRUE))}
```

####False Negative Rate
```{r, warning=FALSE, message=FALSE, error=FALSE, eval=TRUE}
False_Negative_Rate <- function(y,y_hat){
  return(sum(y == 1 & y_hat == 0, na.rm=TRUE) / sum(y == 1 & y_hat != "NA", na.rm=TRUE))}
```

####Sensitivity (True Positive Rate)
```{r, warning=FALSE, message=FALSE, error=FALSE, eval=TRUE}
Sensitivity <- function(y,y_hat){
  return(True_Positive_Rate(y,y_hat))}
```

####Specificity (True Negative Rate)
```{r, warning=FALSE, message=FALSE, error=FALSE, eval=TRUE}
Specificity <- function(y,y_hat){
  return(True_Negative_Rate(y,y_hat))}
```

####Recall (True Positive Rate)
```{r, warning=FALSE, message=FALSE, error=FALSE, eval=TRUE}
Recall <- function(y,y_hat){
  return(True_Positive_Rate(y,y_hat))}
```

####Precision
```{r, warning=FALSE, message=FALSE, error=FALSE, eval=TRUE}
Precision <- function(y,y_hat){
  return(sum(y == 1 & y_hat == 1, na.rm=TRUE) / sum(y_hat == 1 & y != "NA", na.rm=TRUE))}
```

####Prevalence
```{r, warning=FALSE, message=FALSE, error=FALSE, eval=TRUE}
Prevalence <- function(y,y_hat){
  return(sum(y == 1, na.rm=TRUE) / sum(y != "NA", na.rm=TRUE))}
```

####Accuracy
```{r, warning=FALSE, message=FALSE, error=FALSE, eval=TRUE}
Accuracy <- function(y,y_hat){
  return((sum(y == 1 & y_hat == 1, na.rm=TRUE) + sum(y == 0 & y_hat == 0, na.rm=TRUE)) / sum(y != "NA" & y_hat != "NA", na.rm=TRUE))}
```

####Kappa
```{r, warning=FALSE, message=FALSE, error=FALSE, eval=TRUE}
Kappa <- function(y,y_hat){
  return(1-((1-Accuracy(y,y_hat))/
              (1-(((sum(y == 1 & y_hat != "NA", na.rm=TRUE) / sum(y != "NA" & y_hat != "NA", na.rm=TRUE))*(sum(y_hat == 1 & y != "NA", na.rm=TRUE) / sum(y != "NA" & y_hat != "NA", na.rm=TRUE)))+
                  ((sum(y == 0 & y_hat != "NA", na.rm=TRUE) / sum(y != "NA" & y_hat != "NA", na.rm=TRUE))*(sum(y_hat == 0 & y != "NA", na.rm=TRUE) / sum(y != "NA" & y_hat != "NA", na.rm=TRUE)))))))}
```



## Usage 
Use the function(s) above to calculate the metrics for your late arrival model that
you created last week. 

####Load dataset and declare variables y and y_hat
```{r, warning=FALSE, message=FALSE, error=FALSE, eval=TRUE, include=FALSE}
delays <- read.csv("all.csv")
delays <- tbl_df(delays)

delaystrain <- sample_n(delays,10000)
delaystrain

delaystrain$delay22[delaystrain$arr_delay >= 22] <- 1
delaystrain$delay22[delaystrain$arr_delay < 22] <- 0

delaylogistic13 <- glm(formula = delay22 ~ dep_delay + month + air_time + distance + dep_time + humid + temp + pressure + lat.y + wind_gust + carrier, family = binomial, data = delaystrain)
summary(delaylogistic13)

delaystest <- sample_n(delays,10000)
delaystest$delay22[delaystest$arr_delay >= 22] <- 1
delaystest$delay22[delaystest$arr_delay < 22] <- 0
delaystest

prob <- predict(delaylogistic13, delaystest, type="response")
logit.pred <- factor(prob>.5, levels=c(FALSE,TRUE), labels=c(0,1))
logit.pref <- table(delaystest$delay22, logit.pred, dnn=c("Actual", "Predicted"))
logit.pref

delaystest$logit[logit.pred == 0] <- 0
delaystest$logit[logit.pred == 1] <- 1


y <- delaystest[,50]
y

y_hat <- delaystest[,51]
y_hat
```

####True Positive Rate
```{r, warning=FALSE, message=FALSE, error=FALSE, eval=TRUE}
True_Positive_Rate(y,y_hat)
```

####False Positive Rate
```{r, warning=FALSE, message=FALSE, error=FALSE, eval=TRUE}
False_Positive_Rate(y,y_hat)
```

####True Negative Rate
```{r, warning=FALSE, message=FALSE, error=FALSE, eval=TRUE}
True_Negative_Rate(y,y_hat)
```

####False Negative Rate
```{r, warning=FALSE, message=FALSE, error=FALSE, eval=TRUE}
False_Negative_Rate(y,y_hat)
```

####Sensitivity
```{r, warning=FALSE, message=FALSE, error=FALSE, eval=TRUE}
Sensitivity(y,y_hat)
```

####Specificity
```{r, warning=FALSE, message=FALSE, error=FALSE, eval=TRUE}
Specificity(y,y_hat)
```

####Recall
```{r, warning=FALSE, message=FALSE, error=FALSE, eval=TRUE}
Recall(y,y_hat)
```

####Precision
```{r, warning=FALSE, message=FALSE, error=FALSE, eval=TRUE}
Precision(y,y_hat)
```

####Prevalence
```{r, warning=FALSE, message=FALSE, error=FALSE, eval=TRUE}
Prevalence(y,y_hat)
```

####Accuracy
```{r, warning=FALSE, message=FALSE, error=FALSE, eval=TRUE}
Accuracy(y,y_hat)
```

####Kappa
```{r, warning=FALSE, message=FALSE, error=FALSE, eval=TRUE}
Kappa(y,y_hat)
```

