---
title: "01-exercises"
author: "Christopher Brown"
date: "April 6, 2016"
output: html_document
---


## Exercise 1:

Write a function (f) that takes a vector of numbers, **x**, and returns a vector of numbers such that each element containing the product of every element of x except the element of x with the same index. 


### Solution 

```{r}
f <- function(x) {
  
x <- c( 1, 5, 2, 8 )

z <- c(cumprod(x[-1])[3], cumprod(x[-2])[3], cumprod(x[-3])[3], cumprod(x[-4])[3])
z

}

f(x)

```
    



## Exercise 2

Write a function f(x) to accept an integer vector, and returns a vector with those numbers except for: 
   multiples of 3 replaced by "Fizz"? 
   multiples of 5 replaced by "Buzz" 
   multiples of 3 and 5 replaced by "FizzBuzz""
   

### Solution

```{r}
f <- function(x) {

x <- 1:20  
  
f <- seq(3,20,3)
b <- seq(5,20,5)
fizzbuzz <- f[f %in% b] 
fizz <- f[!f %in% fizzbuzz]
buzz <- b[!b %in% fizzbuzz]
x[fizz] <- "Fizz"
x[buzz] <- "Buzz"
x[fizzbuzz] <- "FizzBuzz"
cat(x)
x

}

f(1:20)

```
